The genomic impacts of drift and selection for hybrid performance in maize

### 2015-07-03 methods.tex
supp figures, tables
working on response

We called 48,919 SNPs on the Illumina platform from the MaizeSNP50_B.egt cluster file. Genotypes with quality scores of 50 or less were recoded as missing data. 
Three plants were removed from the data due to an excess of missing data (the derived line B10, a cycle 0 plant from BSSS, and a cycle 8 plant from BSSS). 
In addition, BSCB1 plant 31 from cycle 4 appeared switched with plant 31 from cycle 8 based on our principal component analysis (PCA), so we switched the labels for these two genotypes to correct the mistake. 
To avoid structure among the missing data, we removed any SNP that was coded as missing in more than 3 plants in either group of founders or any group of plants from a particular cycle and population. 
Preliminary analysis by PCA and heatmap plots of distance matrices revealed two additional likely mix-ups. 
Plant 23 from BSSS cycle 8 was a clear outlier from the BSSS population as a whole and plant 2 from BSSS cycle 0 is likely a mislabeled plant from cycle 16. Since there was no evidence suggesting when mis-labelings occurred, each of these plants was removed from the analysis.

@@

Only a portion of the SNPs on the array had genetically mapped positions. 
‘Unmapped’ SNPs (those with a physical position but no genetic position) bounded by two mapped markers were moved along with their mapped neighbors if both anchoring mapped markers were also moved. 
However, it is unclear whether unmapped SNPs just outside of these anchors should be kept in place or moved along with the adjacent SNPs. 
Since most inversions were small relative to the genetic map (and would therefore still fall in the same window of a sliding window analysis), these SNPs were left in place. 
However, markers bordering translocations were removed to ensure there were no markers mapped to the incorrect linkage group. 
Unmapped SNPs were then assigned a genetic position by linear interpolation of genetic vs. physical distance using the approx() function in R \citep{rteam}. 
The IBM genetic map distances were then converted to single-meiosis map distances using the formulae of \citet{winkler2003determination}. 
Finally, SNPs located at physical positions outside of those bounded by the genetic map (such as the telomeres) were assigned the genetic position of their nearest mapped neighbor. \jri{so everything beyond 0 gets a 0 position on the map instead of negative?}
Since moved segments were arbitrarily joined 10 kb from their nearest genetic neighbor, we acknowledge that the physical positions of these markers are only estimates. 
However, the estimated junctions are small relative to the genetic windows used for our analysis. 
The final map used is provided as supplemental material (Supplementary File S1).

>>>>

We called 48,919 SNPs on the Illumina platform from the MaizeSNP50\_B.egt cluster file. Genotypes with quality scores of 50 or less were recoded as missing data. 
Three plants were removed from the data due to an excess of missing data (the derived line B10, a cycle 0 plant from BSSS, and a cycle 8 plant from BSSS). 
In addition, BSCB1 plant 31 from cycle 4 appeared switched with plant 31 from cycle 8 based on our principal component analysis (PCA), so we switched the labels for these two genotypes to correct the mistake. 
To avoid structure among the missing data, we removed any SNP that was coded as missing in more than 3 plants in either group of founders or any group of plants from a particular cycle and population. 
Preliminary analysis by PCA and heatmap plots of distance matrices revealed two additional likely mix-ups. 
Plant 23 from BSSS cycle 8 was a clear outlier from the BSSS population as a whole and plant 2 from BSSS cycle 0 is likely a mislabeled plant from cycle 16. Since there was no evidence suggesting when mis-labelings occurred, each of these plants was removed from the analysis.

@@

Only a portion of the SNPs on the array had genetically mapped positions. 
‘Unmapped’ SNPs (those with a physical position but no genetic position) bounded by two mapped markers were moved along with their mapped neighbors if both anchoring mapped markers were also moved. 
However, it is unclear whether unmapped SNPs just outside of these anchors should be kept in place or moved along with the adjacent SNPs. 
Since most inversions were small relative to the genetic map (and would therefore still fall in the same window of a sliding window analysis), these SNPs were left in place. 
However, markers bordering translocations were removed to ensure there were no markers mapped to the incorrect linkage group. 
Unmapped SNPs were then assigned a genetic position by linear interpolation of genetic vs. physical distance using the approx() function in R \citep{rteam}. 
The IBM genetic map distances were then converted to single-meiosis map distances using the formulae of \citet{winkler2003determination}. 
Finally, SNPs located at physical positions outside of those bounded by the genetic map (such as the telomeres) were assigned the genetic position of their nearest mapped neighbor. 
Since moved segments were arbitrarily joined 10 kb from their nearest genetic neighbor, we acknowledge that the physical positions of these markers are only estimates. 
However, the estimated junctions are small relative to the genetic windows used for our analysis. 
The final map used is provided as supplemental material (Supplementary File S1).

### 2015-07-03 results.tex
working on response

Since the population size of the Iowa RRS is small (10-20), many biallelic SNPs should fix by chance regardless of their starting minor allele frequencies. The deviation from simulation arises not from changes in allele frequencies per se, but rather from the fixation of linked markers across larger than expected genetic distances. 
The validity of significance cutoffs therefore depend on the accuracy of our genetic map, and the maize genetic map is known to vary among genetic backgrounds across short genetic distances (< 5cM) (McMullen et al., 2009). \jri{need to cite either \citep{bauer2013intraspecific} who shows there's lots of variation or maybe also \citep{rodgers2015recombination} who argues it's stable (but stable vs. variable is somewhat subjective}
Deviations of the observed and simulated results could be due to selection, variation / inaccuracy in the genetic map, or a combination of these factors.

>>>>

Since the population size of the Iowa RRS is small (10-20), many biallelic SNPs should fix by chance regardless of their starting minor allele frequencies. The deviation from simulation arises not from changes in allele frequencies per se, but rather from the fixation of linked markers across larger than expected genetic distances. 
The validity of significance cutoffs therefore depend on the accuracy of our genetic map, and the maize genetic map is known to vary among genetic backgrounds across short genetic distances ($<5$cM) (McMullen et al., 2009). \jri{need to cite either \citep{bauer2013intraspecific} who shows there's lots of variation or maybe also \citep{rodgers2015recombination} who argues it's stable (but stable vs. variable is somewhat subjective}
Deviations of the observed and simulated results could be due to selection, variation / inaccuracy in the genetic map, or a combination of these factors.

### 2015-07-03 sup1.tex
working on response

\begin{figure}[tb]   
  \begin{center}
   \includegraphics[width=\linewidth]{figs1_jri}
   \caption{Heterozygosity in each population, observed vs. simulated data. Heterozygosity was calculated as the average across all markers genome-wide  in the BSSS (A)  and BSCB1 (B) populations. The observed data is marked by the red line, the simulations by gray dots, and the median of the simulations by a green dot. } 
    \label{fig:s1}
  \end{center}
\end{figure}

\pagebreak

\begin{figure*}[tb]   
  \begin{center}
\includegraphics[width=\linewidth]{figs2_jri}
   \caption{Heterozygosity ($H$) at cycle 16 across all ten chromosomes in each population, calculated on 15-marker sliding windows with 5 marker steps as in Figure 3. Heterozygosity values in BSSS (blue dots) and BSCB1 (red dots) are superimposed in one panel. 2 cM windows of low heterozygosity that were reproduced in less than 10 of 10,000 simulations (P<0.001) are shaded in light blue (BSSS) or pink (BSCB). At this cutoff, there is no overlap between highlighted windows in the two populations. (A) Physical map. (B) Genetic map. } 
    \label{fig:s2}
  \end{center}
\end{figure*}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE
\begin{figure*}   
  \begin{center}
   \vspace{-0mm}
   \includegraphics[\linewidth]{fig3}
   %\renewcommand{\baselinestretch}{0.9}
   \vspace{-3mm}
   \caption{ Heterozygosity ($H$) at cycle16 across all ten chromosomes in each population.  $H$ is calculated on 15-marker sliding windows with 5 marker steps. Each point is plotted at the midpoint of the 15-marker window. 
} 
\vspace{-6mm}
    \label{fig:s2}
  \end{center}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE
